---
title: "ProteinOrtho Abundance Graph"
author: "Nikeisha Caruana"
date: "11 January 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(stringr)
library(dplyr)
library(splitstackshape)
```


The code below merges two species ProteinGroups.txt files created by MaxQuant, with a ProteinOrtho file from the same two species in order to visualize the abundance of the orthologous proteins between the two species.  

Read in protein orthofiles 

```{r read, include = TRUE}
proteinortho <-  read.csv("../raw_data_shared/proteinortho_data/slimeglyco_comparison.proteinortho", sep='\t', header = TRUE)
proteingroups_SL <- read.csv("../raw_data_shared/proteinGroups_PNGaseFSL.txt",sep='\t', header = TRUE)
proteingroups_SA <- read.csv("../raw_data_shared/proteinGroups_SA.txt",sep='\t', header = TRUE)
```

Expand protein groups so we have one row per protein.
```{r expand function, include = TRUE}
expand_groups <- function(protein_data){
  group_list = lapply(1:nrow(protein_data),function(ri){
    prot_ids = unlist(str_split(protein_data[ri,]$Protein.IDs,pattern = ";"))
    prot_ids = str_match(prot_ids,"lcl\\|(.*)")[,2]
    prot_ids = str_replace(prot_ids,"_m\\.",replacement = "\\|m\\.")
    other_cols = protein_data[rep(ri,length(prot_ids)),]
    data.frame(prot_ids,other_cols)
  })
  do.call(rbind,group_list)
}

```

Expand and clean up proteingroups files

```{r expand and clean, include = TRUE}
expanded_SA <- expand_groups(proteingroups_SA)
expanded_SA$prot_ids <- str_replace(expanded_SA$prot_ids,"\\|m\\..*", "")
expanded_SL <- expand_groups(proteingroups_SL)
expanded_SL$prot_ids <- str_match(expanded_SL$prot_ids,"TRINITY_DN[0-9]*_c[0-9]_g[0-9]_i[0-9]")

#remove unneeded information - just want iBAQ and prot_ids (for now)
expanded_SA <- subset(expanded_SA, select = c("prot_ids","iBAQ"))
expanded_SL <- subset(expanded_SL, select = c("prot_ids","iBAQ"))

#change naming conventions (in order to be able to merge) 
expanded_SL <- rename(expanded_SL, iBAQSL = iBAQ, protID_SL = prot_ids)
expanded_SA <- rename(expanded_SA, iBAQSA = iBAQ, protID_SA = prot_ids)
proteinortho <- rename(proteinortho, protID_SL = TrinitySepio.fasta, protID_SA = TrinitySepia.fasta)
proteinortho <-unique(cSplit(proteinortho, "protID_SA", ",", "long"))
proteinortho <-unique(cSplit(proteinortho, "protID_SL", ",", "long"))

```

Merge the Abundance information to the proteinortho table.
```{r merge, include = TRUE}

#Add iBAQ from expanded_SL to proteinortho
merged <- merge.data.frame(expanded_SL, proteinortho,by.x = "protID_SL", by.y = "protID_SL")
#Add iBAQ from expanded_SA to proteinortho
merged <- merge.data.frame(expanded_SA, merged,by.x = "protID_SA", by.y = "protID_SA")

```






