---
title: "accumulation_curve_all"
author: "Nikeisha Caruana"
date: "24 April 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
rm(list=ls())

library(ggplot2)
library(dplyr) # useful for data manipulation
library(splitstackshape)
library(tidyr)
library(combinat)
library(data.table)

```

## Accumulation Curve for Glycosylated proteins found in PNGaseF samples'

```{r read in data}
TPP_data = read.table("raw_data/TPPalldata.txt",sep="\t",header=TRUE)
TPP_data = TPP_data[TPP_data$interprophet_prob >= 0.99, ]
TPPpeptide = TPP_data %>% select(protein,contains("experiment_label"))

MQ_data = read.table("raw_data/MQpeptides.txt",sep="\t",header=TRUE,na.strings = "0")
MQpeptide = MQ_data %>% select(Leading.razor.protein,contains("Intensity."))
```


```{r clean up data}
MQ_tall <- MQpeptide %>% 
    gather(sample,int_value,-Leading.razor.protein) %>% 
    filter(!is.na(int_value)) %>% # Removes NAs
    filter(!grepl("CON_",Leading.razor.protein)) %>%
    filter(!grepl("sp",Leading.razor.protein)) %>%# Removes contaminants
    filter(!grepl("decoy",Leading.razor.protein)) #Removes decoys

colnames(MQ_tall)[1] <- "protein"


TPP_tall <- TPPpeptide %>% 
    gather(sample,exp_label,-protein) %>% 
    filter(!is.na(exp_label)) %>% # Removes NAs
    filter(!grepl("decoy",protein)) %>%#Removes decoys
    unite(experiment, c(sample,exp_label))

colnames(TPP_tall)[2] <- "sample"

peptide_all <- merge(MQ_tall, TPP_tall, all = TRUE)
```


***IRA***

Up until this point I think its okay. I think the problem in the data is my understanding and manipulation of
the functions. For MQ I think I have a total of 2275 peptides and for the TPP I have 3,500-4,500 (approximately) peptides. What Im getting after the permutations is too low. 

***IRA***

```{r}
distinct_proteins_in_samples <- function(samples){
    peptide_all %>% 
    filter(sample %in% samples) %>% 
    select(protein) %>% 
    distinct()
}

new_proteins <- function(samples_1,samples_2){
  prots_samples1 <- distinct_proteins_in_samples(samples_1)
  prots_samples2 <- distinct_proteins_in_samples(samples_2)
  
  setdiff(prots_samples2,prots_samples1)
}

accumulation <- function(samples){
  n1 <- nrow(distinct_proteins_in_samples(samples[1]))
  n_prots <- sapply(2:length(samples),function(i){
    nrow(new_proteins(samples[1:(i-1)],samples[i]))
  })
  c(n1,n_prots)
}
```

```{r}
MQ_P <- colnames(MQpeptide)[c(3,5,7,9)]
MQ_NP <- colnames(MQpeptide)[c(2,4,6,8)]
TPP_NP <- c("experiment_label_1","experiment_label_2","experiment_label_3","experiment_label_4")
TPP_P <- c("experiment_label_1P","experiment_label_2P","experiment_label_3P","experiment_label_4P")

acc_MQNP <- sapply(permn(MQ_NP),accumulation)
acc_MQP <- sapply(permn(MQ_P),accumulation)
acc_TPPNP <- sapply(permn(TPP_NP),accumulation)
acc_TPPP <- sapply(permn(TPP_P),accumulation)

```

***IRA***
When I run the permn functions the results are too low for the amount of peptides.
***IRA***



```{r}
plot_dataMQNP <- as.data.frame(t(apply(acc_MQNP,2,cumsum))) %>% gather(sample_num,total_proteins)
plot_dataMQP <- as.data.frame(t(apply(acc_MQP,2,cumsum))) %>% gather(sample_num,total_proteins)
plot_dataTPPNP <- as.data.frame(t(apply(acc_TPPNP,2,cumsum))) %>% gather(sample_num,total_proteins)
plot_dataTPPP <- as.data.frame(t(apply(acc_TPPP,2,cumsum))) %>% gather(sample_num,total_proteins)


plot_dataMQNP2 <- aggregate(total_proteins~sample_num,plot_dataMQNP,mean) 
plot_dataMQP2 <- aggregate(total_proteins~sample_num,plot_dataMQP,mean)
plot_dataTPPNP2 <- aggregate(total_proteins~sample_num,plot_dataTPPNP,mean)
plot_dataTPPP2 <- aggregate(total_proteins~sample_num,plot_dataTPPP,mean) 

plot_dataMQNP2$sample_num <- gsub("V", "",plot_dataMQNP2$sample_num)
plot_dataMQP2$sample_num <-gsub("V", "",plot_dataMQP2$sample_num)
plot_dataTPPNP2$sample_num <- gsub("V", "",plot_dataTPPNP2$sample_num)
plot_dataTPPP2$sample_num<- gsub("V", "",plot_dataTPPP2$sample_num)

 ```

```{r plotting, echo=FALSE}

ggplot() + 
  geom_point(size=2, data = plot_dataMQNP2, aes(x=sample_num,y=total_proteins, colour = "MQ Non-PNGase F")) +
  geom_point(size=2, data = plot_dataMQP2, aes(x=sample_num,y=total_proteins, colour = "MQ PNGase F")) + 
  geom_point(size=2, data = plot_dataTPPNP2, aes(x=sample_num,y=total_proteins, colour = "TPP Non-PNGase F")) +
  geom_point(size=2, data = plot_dataTPPP2, aes(x=sample_num,y=total_proteins, colour = "TPP PNGase F")) + 
  xlab('Sample Number') + ylab ('Number Total Peptides') + theme(legend.title=element_blank())

```
